Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-27 analysis/paper/hotspots-paper.Rmd
moved eval chunk argument to an "if" conditional statement

```{r load_attie, eval=!file.exists(here("analysis", "data", "derived_data", "expr.rds"))}
download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
```

@@

```{r scatter-2}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

```{r scatter_by_local-2, fig.height = ceiling(nlocal / 2)}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
p <- foo %>%
  ggplot() +
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") +
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  xlab("Pleiotropy LOD") +
  ylab("Mediation LOD difference") +
  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```

>>>>

```{r load_attie}
if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
}
```

@@

```{r scatter-2}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol, local = local_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
plotly::ggplotly(p)
```

```{r scatter_by_local-2, fig.height = ceiling(nlocal / 2)}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
p <- foo %>%
  ggplot() +
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") +
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) +
  xlab("Pleiotropy LOD") +
  ylab("Mediation LOD difference") +
  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
```
